---
title: "ANCOM-BC_16S_VOCPCA"
output: html_document
date: "2025-03-04"
---

#load

```{r}
library(readxl)
library(phyloseq)
library(SummarizedExperiment)
library(microbiome)
library(ANCOMBC)
library(tidyverse)
```

#path to raw-data directory (change "path" to your raw-data directory)

```{r}
#create "path" file, and define "path" as your raw-data directory(character vector) in the file
#"path" is raw-data directory, and "path_out" is output file directory

source("../scripts/path.R")
source("../R/plot_top_taxa.R")
```

#theme

```{r}
#theme settings
mytheme <-theme(text = element_text(color = "black"), 
                axis.text = element_text(color = "black"), 
                legend.key.size = unit(8, "pt"), 
                legend.spacing.x = unit(0, "mm"))
```

#read data

```{r}
#the microbiome outlier (AZ_040_1, based on bray) is excluded in advance
#rarefied-ASV table
df_rarefied <- read.csv(paste0(path_out, "/231202_Sugi_phyllosphere_June_ITS/processed_data/BVOC/df_rarefied_ITS_BVOC.csv"), row.names = 1)
rownames(df_rarefied) <- formatC(as.numeric(rownames(df_rarefied)), width = 4, flag = "0") 
#add 0 to the head of rownames

#tax_table
tax_table <- read.csv(paste0(path_out, "/231202_Sugi_phyllosphere_June_ITS/processed_data/tax_table2_ITS.csv"), row.names = 1)
tax_table <- as.matrix(tax_table)

#phylogenetic tree
load(paste0(path_out, "/231202_Sugi_phyllosphere_June_ITS/processed_data/trees.Rdata"))

#metadata
metadata <- read.csv(paste0(path_out, "/231202_Sugi_phyllosphere_June_ITS/processed_data/BVOC/metadata_ITS_BVOC.csv"), row.names = 1)
rownames(metadata) <- formatC(as.numeric(rownames(metadata)), width = 4, flag = "0")

metadata$cloneID <- as.factor(metadata$cloneID)
metadata$population <- as.factor(metadata$population)
metadata$site <- as.factor(metadata$site)
metadata$date <- as.factor(metadata$date)
```

```{r}
#read VOC data
load(file = paste0(path_out, "/data/Sugi/BVOC/exp_Terpene_2209to2408_LODrev_nolod.Rdata"))
load(file = paste0(path_out, "/data/Sugi/BVOC/exp_Terpene_2209to2408_LODrev_log.Rdata"))
#exp_Ohta_nolod is the data in which LOD substituted with 0
#exp_Ohta_log is the log-transformed data

#extract the data for June 2023
BVOC_2306 <- exp_Ohta_nolod %>% subset(Month == "2306")
BVOC_2306_log <- exp_Ohta_log %>% subset(Month == "2306")
#microbial samples for AJ_002_1 and AZ_004_1 were not collected, so remove them
#"AZ_040_1" is also removed because it is an outlier based on bray distance
BVOC_2306 <- BVOC_2306 %>% subset(!(treeID %in% c("AJ_002_01", "AZ_004_01", "AZ_040_1")))
BVOC_2306_log <- BVOC_2306_log %>% subset(!(treeID %in% c("AJ_002_01", "AZ_004_01", "AZ_040_1")))

#match the rownames with those of the microbial data
#feature to match the rownames
rowData(BVOC_2306) <- data.frame(rowData(BVOC_2306)) %>%
  mutate(samplename = paste(subpop, IDmid, str_sub(cloneID, 2), sep = "_"))
rowData(BVOC_2306_log) <- data.frame(rowData(BVOC_2306_log)) %>%
  mutate(samplename = paste(subpop, IDmid, str_sub(cloneID, 2), sep = "_"))

rownames(BVOC_2306) <- rownames(metadata[match(rowData(BVOC_2306)$samplename, 
                                               metadata$samplename), 
                                         ])
rownames(BVOC_2306_log) <- rownames(metadata[match(rowData(BVOC_2306_log)$samplename,
                                                   metadata$samplename),
                                             ])

#Longifolene is not detected in the June 2023 samples, so remove it
BVOC_2306 <- BVOC_2306[, -which(colnames(BVOC_2306) == "Longifolene")]
BVOC_2306_log <- BVOC_2306_log[, -which(colnames(BVOC_2306_log) == "Longifolene")]

#remove beta-Caryophyllene from the dataframe
bCary <- which("beta-Caryophyllene" == colnames(BVOC_2306_log)) #the column number of beta-Caryophyllene
BVOC_2306_log_rm <- assay(BVOC_2306_log)[, -bCary] #remove beta-Caryophyllene
```

```{r}
#calculate log-transformed total BVOC- and MTs emission
BVOC_2306_total_log <- data.frame(total = log(rowSums(assay(BVOC_2306)), base = 10)) %>%
  rownames_to_column(var = "sampleID")
MT_2306_log <- data.frame(MTs = log(rowSums(assay(BVOC_2306[, BVOC_2306$mol_weight == "Mono"])) + 1, base = 10)) %>%
  rownames_to_column(var = "sampleID")
SQT_2306_log <- data.frame(SQT = log(rowSums(assay(BVOC_2306[, BVOC_2306$mol_weight == "Sesqui"])) + 1, base = 10)) %>%
  rownames_to_column(var = "sampleID")
DT_2306_log <- data.frame(DT = log(rowSums(assay(BVOC_2306[, BVOC_2306$mol_weight == "Di"])) + 1, base = 10)) %>%
  rownames_to_column(var = "sampleID")

#merge BVOC data
all_2306_log <- BVOC_2306_log_rm %>%
  rownames_to_column(var = "sampleID") %>%
  full_join(BVOC_2306_total_log, by = "sampleID") %>%
  full_join(MT_2306_log, by = "sampleID") %>%
  full_join(SQT_2306_log, by = "sampleID") %>%
  full_join(DT_2306_log, by = "sampleID")
```

#phyloseq object

```{r}
ps_ITS_BVOC <- phyloseq(otu_table(t(df_rarefied), taxa_are_rows = TRUE),
                            tax_table(tax_table),
                            sample_data(metadata),
                            phyloseq::phy_tree(tree_rarefied2$data))
```

#edit sample_data

```{r}
#add VOC emission profile to 16S sample_data
samdata_ITS <- data.frame(sample_data(ps_ITS_BVOC)) %>%
  rownames_to_column(var = "sampleID")

samdata_new <- left_join(samdata_ITS, 
                         all_2306_log,
                         by = "sampleID") %>%
  column_to_rownames(var = "sampleID")

sample_data(ps_ITS_BVOC) <- sample_data(samdata_new)
```

#plot bar

```{r}
ps_ITS_bF <- subset_samples(ps_ITS_BVOC, 
                           `beta.Farnesene` > 0) #ps without the samples which emitted no detectable beta-Farnesene
ps_ITS_entK <- subset_samples(ps_ITS_BVOC, 
                            `ent.Kaurene` > 0) #ps without the samples which emitted no detectable ent-Kaurene

T_asc <- rownames(data.frame(sample_data(ps_ITS_entK)) %>% arrange(ent.Kaurene))
bar_ITS <- plot_top_taxa(ps_ITS_entK,
                         n_taxa = 20,
                         taxonomic_level = "Class", 
                         transform = "relative", 
                         show_unknown = FALSE, 
                         sample_order = T_asc)

bar_ITS$plot <- bar_ITS$plot +
  theme_classic(base_size = 12) + mytheme +
  xlab("Low - ent-Kaurene - High")
  #facet_wrap(~population, 
  #           scale = "free_x")
show(bar_ITS$plot)
```

#ANCOM-BC

```{r}
load(paste0(path_out, "/231202_Sugi_phyllosphere_June_ITS/analysis/ancombc_ITS_BVOC2306.Rdata"))
```

```{r,eval=FALSE}
ancombc_ITS_total <- ancombc2(data = ps_ITS_BVOC,
                              tax_level = "Class",
                              fix_formula = "total",
                              p_adj_method = "BH",
                              prv_cut = 0.1, #taxa with less than 10% occurrence are cutted
                              lib_cut = 0, #library size thresholds
                              s0_perc = 0.05, #small count to add as pseudo count. % of 
                              #group = "population", #group to detect structural zeros
                              struc_zero = FALSE,
                              neg_lb = FALSE,
                              alpha = 0.05, #significance
                              global = FALSE, #global test
                              verbose = TRUE)
```

#results

```{r}
show(ancombc_ITS_total$res)
```

```{r,eval=FALSE}
ancombc_ITS_MTs <- ancombc2(data = ps_ITS_BVOC,
                            tax_level = "Class",
                            fix_formula = "MTs",
                            p_adj_method = "BH",
                            prv_cut = 0.1, #taxa with less than 10% occurrence are cutted
                            lib_cut = 0, #library size thresholds
                            s0_perc = 0.05, #small count to add as pseudo count. % of
                            #group = "population", #group to detect structural zeros
                            struc_zero = FALSE,
                            neg_lb = FALSE,
                            alpha = 0.05, #significance
                            global = FALSE, #global test
                            verbose = TRUE)
```

```{r}
show(ancombc_ITS_MTs$res)
```

```{r,eval=FALSE}
#subset samples with beta-Farnesene > 0
ps_ITS_bF <- subset_samples(ps_ITS_BVOC, 
                           `beta.Farnesene` > 0)
#when LOD values are not removed, no taxa are detected as DA

ancombc_ITS_bF <- ancombc2(data = ps_ITS_BVOC,
                           tax_level = "Class",
                           fix_formula = "beta.Farnesene",
                           p_adj_method = "BH",
                           prv_cut = 0.1, #taxa with less than 10% occurrence are cutted
                           lib_cut = 0, #library size thresholds
                           s0_perc = 0.05, #small count to add as pseudo count. % of
                           #group = "population", #group to detect structural zeros
                           struc_zero = FALSE,
                           neg_lb = FALSE,
                           alpha = 0.05, #significance
                           global = FALSE, #global test
                           verbose = TRUE)
```

```{r}
show(ancombc_ITS_bF$res)
```

```{r,eval=FALSE}
#subset samples with ent-Kaurene > 0
ps_ITS_entK <- subset_samples(ps_ITS_BVOC, 
                            `ent.Kaurene` > 0)
ps_ITS_entK <- prune_taxa(taxa_sums(otu_table(ps_ITS_entK) > 0), ps_ITS_entK) #remove the taxa with 0 sum

ancombc_ITS_entK <- ancombc2(data = ps_ITS_entK,
                           tax_level = "Class",
                           fix_formula = "ent.Kaurene",
                           p_adj_method = "BH",
                           prv_cut = 0.1, #taxa with less than 10% occurrence are cutted
                           lib_cut = 0, #library size thresholds
                           s0_perc = 0.05, #small count to add as pseudo count. % of
                           #group = "population", #group to detect structural zeros
                           struc_zero = FALSE,
                           neg_lb = FALSE,
                           alpha = 0.05, #significance
                           global = FALSE, #global test
                           verbose = TRUE)
```

```{r}
show(ancombc_ITS_entK$res)
```

<tr>**DAs were detected mostly for the rare taxa (with small relative abundances and occurrences)**
<tr>**Eurotiomycetes was negatively correlated with the ent-Kaurene emission, although it didn't pass SA and p-adjustment.**

##save the results

```{r}
save(ancombc_ITS_total, 
     ancombc_ITS_MTs, 
     ancombc_ITS_bF, 
     ancombc_ITS_entK, 
     file = paste0(path_out, "/231202_Sugi_phyllosphere_June_ITS/analysis/ancombc_ITS_BVOC2306.Rdata"))
```

#lfc plot
##230824

```{r,eval=FALSE}
#AZ to AJ
df_230824 <- ancombc_16Sfilt$res %>%
    dplyr::select(taxon, ends_with("date230824")) #select stats related with date230824
df_fig_230824 <- df_230824 %>%
    dplyr::filter(diff_date230824 == 1) %>% #differentially abundant in 230824 compared to AJ
    dplyr::arrange(desc(lfc_date230824)) %>% #log fold changes in date230824 to AJ
    dplyr::mutate(direct = ifelse(lfc_date230824 > 0, "Positive LFC", "Negative LFC"),
                  color = ifelse(passed_ss_date230824 == 1, "aquamarine3", "black"))
#shorten the taxon names for uncultured taxa
df_fig_230824$taxon <- sapply(lapply(str_split(df_fig_230824$taxon, "_"),
                                     tail, 2),
                              paste0, collapse = "_")
df_fig_230824$taxon = factor(df_fig_230824$taxon, levels = df_fig_230824$taxon)
df_fig_230824$direct = factor(df_fig_230824$direct, 
                           levels = c("Positive LFC", "Negative LFC"))

fig_230824 = df_fig_230824 %>%
    ggplot(aes(x = taxon, y = lfc_date230824, fill = direct)) + 
    geom_bar(stat = "identity", width = 0.7, color = "black", 
             position = position_dodge(width = 0.4)) +
    geom_errorbar(aes(ymin = lfc_date230824 - se_date230824, ymax = lfc_date230824 + se_date230824), 
                  width = 0.2, position = position_dodge(0.05), color = "black") + 
    labs(x = NULL, y = "Log fold change", 
         title = "Log fold changes in 230824 compared to 230602") + 
    scale_fill_discrete(name = NULL) +
    scale_color_discrete(name = NULL) +
    theme_classic(base_size = 12) + mytheme +
    theme(plot.title = element_text(hjust = 0.5),
          panel.grid.minor.y = element_blank(),
          axis.text.x = element_text(angle = 60, hjust = 1,
                                     color = df_fig_230824$color))
fig_230824

ggsave(fig_230824, 
       file = paste0(path_out, "/figures/16S/ANCOM-BC_LFC_16Sfilt_inVOC_230824_Genus.pdf"), 
       width = 160, height = 120, units = "mm", dpi = 300)
```



```{r}

```

