---
title: "pca_BVOC"
output: html_document
date: "2024-10-10"
---

#load packages

```{r}
library(readxl)
library(phyloseq)
library(SummarizedExperiment)
library(tidyverse)
library(ggtext)
library(RColorBrewer)
library(viridis)
library(dichromat)
library(scales)
library(grid)
library(gridExtra)
#do this only once
#loadfonts()
```

#path to raw-data directory (change "path" to your raw-data directory)

```{r}
#create "path" file, and define "path" as your raw-data directory(character vector) in the file
#"path" is raw-data directory, and "path_out" is output file directory

source("../scripts/path.R")
```

#source codes

```{r}
source("../R/label.R")
source("../R/greek_as_alphabet.R")
source("../R/plot_top_taxa.R")
```

#read processed data

```{r import}
ter_total_meta <- read.csv(paste0(path_outall, "/data/Sugi/BVOC/Terpene_metadata_2007to2408.csv"))
ter_total_meta <- column_to_rownames(ter_total_meta, var = "X")

#population metadata
meta_pop <- read_excel(paste0(path_outall, "/data/Sugi/metadata_2306.xlsx"), sheet = "location")

#result of PCA on climatic variables
load(paste0(path_outall, "/data/Sugi/pca_climate.Rdata"))
df_pcaclim <- data.frame(pca_clim$x)
colnames(df_pcaclim) <- paste0(colnames(df_pcaclim), "_clim")

#merge population metadata and PCA result
meta_pop <- cbind(meta_pop, df_pcaclim)

#change subpopulation ID to population ID
ter_total_meta <-  ter_total_meta %>%
  mutate(population = str_replace_all(Population,
                           pattern = c(AT = "SM", HN = "YK", ST = "YK", LD = "YK")))

#add the result of the climatic PCA to the metadata
meta_merge <- left_join(ter_total_meta, meta_pop, by = "population") %>%
  dplyr::select(-c(mat, apr, cvmpr, sdmmt, hurs, pet, sub0pr, PC3_clim, PC4_clim, PC5_clim, PC6_clim, PC7_clim))

rownames(meta_merge) <- rownames(ter_total_meta)
```

```{r import}
#BVOC data
load(paste0(path_outall, "/data/Sugi/BVOC/Terpene_emitted_2007to2408.Rdata"))
load(paste0(path_outall, "/data/Sugi/BVOC/Terpene_stored_K2007.Rdata"))
ter_molwei <- read.csv(paste0(path_outall, "/data/Sugi/BVOC/Terpene_molcular-weight.csv"))
ter_molwei <- column_to_rownames(ter_molwei, var = "X")

#re-order terpene emission data based on molcular weight data
ter_total <- ter_total %>%
  dplyr::select((rownames(ter_molwei)))
```

##assign colors

```{r}
#assign colors to each molcular weight
##13 monoterpene, 6 sesquiterpene, 2 diterpene
col_mono <- viridis::magma(16)[4:16]
col_sesqui <- viridis(10)[5:10]
col_di <- viridis(10)[3:4]
col_ter <- c(col_mono, col_sesqui, col_di)

names(col_ter) <- ter_molwei$Terpene

ter_molwei$col <- col_ter
```

```{r}
#check for visibility to dichromats
par(mfrow = c(4, 1), mar = c(0, 5, 0, 0))
image(1:17, 1, matrix(1:17, 17),
      col = col_ter, axes = FALSE,
      xlab = "", ylab = "Original",
      cex.lab = 2.5)
CheckCol <- c("deutan", "protan", "tritan")
########

#conduct dichromat commands
for(n in seq(CheckCol)){
  
  GetCol <- dichromat(col_ter, type = CheckCol[n])
  #plot using base::imade commands
  image(1:17, 1, matrix(1:17, 17),
        col = GetCol, axes = FALSE,
        xlab = "", ylab = CheckCol[n],
        cex.lab = 2.5)
}
```

The visibility decreases for dichromats, but they will still be able to distinguish Mono, Sesqui and Di-terpene. 

##phyloseq

```{r}
ps_ter <- phyloseq(otu_table = otu_table(as.matrix(ter_total), taxa_are_rows = FALSE), 
                   tax_table = tax_table(as.matrix(ter_molwei)), 
                   sample_data(meta_merge))

#omit Yoshioka data and beta-Ocimene
ps_Ohta <- prune_samples(sample_data(ps_ter)[["Month"]] != "2007", ps_ter)
ps_Ohta <- subset_taxa(ps_Ohta, Terpene!="beta-Ocimene")
#prune terpenes that are not at all emitted
ps_Ohta <- prune_taxa(taxa_sums(ps_Ohta) > 0, ps_Ohta)

#save(ps_ter, 
#     file = paste0(path_outall, "/data/Sugi/BVOC/phyloseq_Terpene_2007to2408.Rdata"))
#save(ps_Ohta, 
#     file = paste0(path_outall, "/data/Sugi/BVOC/phyloseq_Terpene_2209to2408.Rdata"))
```

```{r}
#divide basal emission rates by molcular weights of the terpenes
##basal emission rate (nmol gdw^-1 h^-1)
ps_ter_mol <- ps_ter

tb_ter <- as.matrix(otu_table(ps_ter_mol))
#extract the molcular weight vector
weight_ter <- as.numeric(as.matrix(tax_table(ps_ter_mol))[, "weight"])
names(weight_ter) <- rownames(tax_table(ps_ter_mol))
#arrange the vector according to the order in the terpene table
weight_ter <- weight_ter[match(colnames(tb_ter), names(weight_ter))]

#If taxa_are_rows = FALSE, transpose the terpene table
#Then, divide it with the molcular weight vector, and re-transpose
tb_ter_mol <- otu_table(t(t(tb_ter) / weight_ter), taxa_are_rows = FALSE)

otu_table(ps_ter_mol) <- tb_ter_mol


#with Ohta phyloseq
ps_Ohta_mol <- ps_Ohta

tb_Ohta <- as.matrix(otu_table(ps_Ohta_mol))
#extract the molcular weight vector
weight_Ohta <- as.numeric(as.matrix(tax_table(ps_Ohta_mol))[, "weight"])
names(weight_Ohta) <- rownames(tax_table(ps_Ohta_mol))
#arrange the vector according to the order in the terpene table
weight_Ohta <- weight_Ohta[match(colnames(tb_Ohta), names(weight_Ohta))]

#If taxa_are_rows = FALSE, transpose the terpene table
#Then, divide it with the molcular weight vector, and re-transpose
tb_Ohta_mol <- otu_table(t(t(tb_Ohta) / weight_Ohta), taxa_are_rows = FALSE)

otu_table(ps_Ohta_mol) <- tb_Ohta_mol

#save(ps_ter_mol, 
#     file = paste0(path_outall, "/data/Sugi/BVOC/phyloseq_Terpene_2007to2408_mol.Rdata"))
#save(ps_Ohta_mol, 
#     file = paste0(path_outall, "/data/Sugi/BVOC/phyloseq_Terpene_2209to2408_mol.Rdata"))
```

#theme settings

```{r}
#theme settings
mytheme <-theme(text = element_text(color = "black"), 
                axis.text = element_text(color = "black"), 
                legend.key.size = unit(8, "pt"), 
                legend.spacing.x = unit(0, "mm"))
```

#Barplot of basal emission rate
##all data

```{r}
#omit NA from phyloseq
##prune beta-Ocimene and Methyl_salicylate from data
ps_ter_pr <- subset_taxa(ps_ter, Terpene!="beta-Ocimene")
ps_ter_pr <- subset_taxa(ps_ter_pr, Terpene!="Methyl_salicylate")
##prune samples with NA
ps_ter_pr <- prune_samples(!is.na(sample_sums(ps_ter_pr)), ps_ter_pr)

#set the color of each terpene
ter_col_all <- as.character(tax_table(ps_ter_pr)[, "col"])
names(ter_col_all) <- as.character(tax_table(ps_ter_pr)[, "Terpene"])

#elong the dataset
ter_all_longer <- as.data.frame(otu_table(ps_ter_pr)) %>%
  mutate(Sample = rownames(.)) %>%
  pivot_longer(cols = -Sample, names_to = "Terpene", values_to = "emission")

#add metadata
meta_all <- data.frame(sample_data(ps_ter_pr)) %>%
  mutate(Sample = rownames(.))

ter_all_longer <-  ter_all_longer %>%
  left_join(meta_all, by = "Sample")
```

```{r}
#depict barplot
p_ter_all <- ggplot(data = ter_all_longer, mapping = aes(x = Sample, y = emission, 
                                                         fill = factor(Terpene, 
                                                                       levels = names(ter_col_all)))) +
  geom_bar(stat = "identity", position = "fill") +
  facet_wrap(~Month, 
             scales = "free_x") +
  theme_classic(base_size = 11) + mytheme +
  scale_fill_manual(values = ter_col_all) +
  scale_y_continuous(labels = label_comma()) +
  labs(fill = "Terpene") +
  ylab("Relative basal emission rate") +
  theme(text = element_text(color = "black"), 
        strip.text = element_text(size = 6),
        legend.title = element_text(size = 8), 
        legend.text = element_markdown(size = 7),
        axis.title.x = element_blank(), 
        axis.text.x = element_text(size = 6, angle = 90))
show(p_ter_all)

ggsave(plot = p_ter_all, filename = paste0(path_out, "/figures/BVOC/BVOC_barrel_2007to2408.pdf"), 
       dpi = 300, width = 160, height = 120, units = "mm")
```

```{r}
###total basal emission rate
p_tertot_all <- plot_bar(ps_ter_pr, fill = "mol_weight") +
  geom_bar(aes(fill=mol_weight), stat="identity", position="stack") +
  scale_fill_viridis_d(labels = c("Diterpene", "Monoterpene", "Sesquiterpene")) +
  scale_y_continuous(labels = label_comma()) +
  theme_classic(base_size = 11) + mytheme +
  facet_wrap(~Month, 
             scales = "free_x") +
  labs(fill = "Terpene") +
  ylab(expression(paste("Basal emission rate (ng ", {gdw^-1}, " ", {h^-1}, ")"))) +
  theme(text = element_text(color = "black"), 
        strip.text = element_text(size = 6),
        legend.title = element_text(size = 8), 
        legend.text = element_markdown(size = 7),
        axis.title.x = element_blank(), 
        axis.text.x = element_text(size = 6, angle = 90))

show(p_tertot_all)

ggsave(plot = p_tertot_all, filename = paste0(path_out, "/figures/BVOC/BVOC_barabs_2007to2408.pdf"), 
       dpi = 300, width = 160, height = 120, units = "mm")
```

The difference in the basal emission rates is not easily understandable since one sample emitted a large amount of terpenes.
Omit it from the bars.

```{r}
hist(sample_sums(ps_ter_pr))
which(sample_sums(ps_ter_pr) > 100000)
#AZ_018 in Kawatabi 23/06 emitted a vast amount of terpenes
#omit this sample from the dataset
ps_ter_pr_omit <- prune_samples(sample_sums(ps_ter_pr) <= 100000, ps_ter_pr)

#arrange samples along with MT ratio
ps_ter_MTratio <- tax_glom(ps_ter_pr_omit, "mol_weight")
ps_ter_MTratio <- transform_sample_counts(ps_ter_MTratio, function(x) x / sum(x))
ter_MTratio <- as.numeric(otu_table(ps_ter_MTratio)[, "Sabinene"])
names(ter_MTratio) <- sample_names(ps_ter_MTratio)
ter_MTratio <- sort(ter_MTratio)

#elong the dataset
ter_omit_longer <- as.data.frame(otu_table(tax_glom(ps_ter_pr_omit, "mol_weight"))) %>%
  mutate(Sample = rownames(.)) %>%
  pivot_longer(cols = -Sample, names_to = "Terpene", values_to = "emission")

#add metadata
meta_omit <- data.frame(sample_data(ps_ter_pr_omit)) %>%
  mutate(Sample = rownames(.))

ter_omit_longer <-  ter_omit_longer %>%
  left_join(meta_omit, by = "Sample")

ter_omit_longer$Sample <- factor(ter_omit_longer$Sample, levels = names(ter_MTratio))
ter_omit_longer$Terpene <- factor(ter_omit_longer$Terpene, levels = c("ent-Kaurene", "Sabinene", "beta-Farnesene"))
```


```{r}
p_tertot_all_omit <- ggplot(data = ter_omit_longer, 
                            mapping = aes(x = Sample, y = emission, 
                                          fill = Terpene)) +
  geom_bar(aes(fill=Terpene), stat="identity", position="stack") +
  scale_fill_viridis_d(labels = c("Diterpene", "Monoterpene", "Sesquiterpene")) +
  scale_y_continuous(labels = label_comma()) +
  theme_classic(base_size = 11) + mytheme +
  xlab("Proportion of monoterpenes") + labs(fill = "Terpene") +
  ylab(expression(paste("Basal emission rate (ng ", {gdw^-1}, " ", {h^-1}, ")"))) +
  theme(text = element_text(color = "black"), 
        strip.text = element_text(size = 6),
        legend.title = element_text(size = 8), 
        legend.text = element_markdown(size = 7),
        axis.text.x = element_blank())

show(p_tertot_all_omit)

ggsave(plot = p_tertot_all_omit, filename = paste0(path_out, "/figures/BVOC/BVOC_barabs_2007to2408_MTorder.pdf"), 
       dpi = 300, width = 160, height = 120, units = "mm")
```

The basal emittion rate of terpenes is constant regardless of the proportion of MTs?

##Ohta data

```{r}
#set the color of each terpene
ter_col_Ohta <- as.character(tax_table(ps_Ohta)[, "col"])
names(ter_col_Ohta) <- as.character(tax_table(ps_Ohta)[, "Terpene"])

#elong the dataset
ter_Ohta_longer <- as.data.frame(otu_table(ps_Ohta)) %>%
  mutate(Sample = rownames(.)) %>%
  pivot_longer(cols = -Sample, names_to = "Terpene", values_to = "emission")

#add metadata
meta_Ohta <- data.frame(sample_data(ps_Ohta)) %>%
  mutate(Sample = rownames(.))

ter_Ohta_longer <-  ter_Ohta_longer %>%
  left_join(meta_Ohta, by = "Sample")
```

```{r}
#depict barplot
p_ter <- ggplot(data = ter_Ohta_longer, mapping = aes(x = Sample, y = emission, 
                                                       fill = factor(Terpene, 
                                                                     levels = names(ter_col_Ohta)))) +
  geom_bar(stat = "identity", position = "fill") +
  facet_wrap(~Month, 
             scales = "free_x") +
  theme_classic(base_size = 11) + mytheme +
  scale_fill_manual(values = ter_col_Ohta) +
  scale_y_continuous(labels = label_comma()) +
  labs(fill = "Terpene") +
  ylab("Relative basal emission rate") +
  theme(text = element_text(color = "black"), 
        strip.text = element_text(size = 6),
        legend.title = element_text(size = 8), 
        legend.text = element_markdown(size = 7),
        axis.title.x = element_blank(), 
        axis.text.x = element_text(size = 6, angle = 90))
show(p_ter)

ggsave(plot = p_ter, filename = paste0(path_out, "/figures/BVOC/BVOC_barrel_2209to2408.pdf"), 
       dpi = 300, width = 160, height = 120, units = "mm")
```

```{r}
###total basal emission rate
p_tertot <- plot_bar(ps_Ohta, fill = "mol_weight") +
  geom_bar(aes(fill=mol_weight), stat="identity", position="stack") +
  scale_fill_viridis_d(labels = c("Diterpene", "Monoterpene", "Sesquiterpene")) +
  scale_y_continuous(labels = label_comma()) +
  theme_classic(base_size = 11) + mytheme +
  facet_wrap(~Month, 
             scales = "free_x") +
  labs(fill = "Terpene") +
  ylab(expression(paste("Basal emission rate (ng ", {gdw^-1}, " ", {h^-1}, ")"))) +
  theme(text = element_text(color = "black"), 
        strip.text = element_text(size = 6),
        legend.title = element_text(size = 8), 
        legend.text = element_markdown(size = 7),
        axis.title.x = element_blank(), 
        axis.text.x = element_text(size = 6, angle = 90))

ggsave(plot = p_tertot, filename = paste0(path_out, "/figures/BVOC/BVOC_barabs_2209to2408.pdf"), 
       dpi = 300, width = 160, height = 120, units = "mm")
```

#Mol base
##all

```{r}
#omit NA from phyloseq
##prune beta-Ocimene and Methyl_salicylate from data
ps_ter_mol_pr <- subset_taxa(ps_ter_mol, Terpene!="beta-Ocimene")
ps_ter_mol_pr <- subset_taxa(ps_ter_mol_pr, Terpene!="Methyl_salicylate")
##prune samples with NA
ps_ter_mol_pr <- prune_samples(!is.na(sample_sums(ps_ter_mol_pr)), ps_ter_mol_pr)

#set the color of each terpene
ter_col_all <- as.character(tax_table(ps_ter_mol_pr)[, "col"])
names(ter_col_all) <- as.character(tax_table(ps_ter_mol_pr)[, "Terpene"])

#elong the dataset
ter_all_mol_longer <- as.data.frame(otu_table(ps_ter_mol_pr)) %>%
  mutate(Sample = rownames(.)) %>%
  pivot_longer(cols = -Sample, names_to = "Terpene", values_to = "emission")

#add metadata
meta_all_mol <- data.frame(sample_data(ps_ter_mol_pr)) %>%
  mutate(Sample = rownames(.))

ter_all_mol_longer <-  ter_all_mol_longer %>%
  left_join(meta_all_mol, by = "Sample")
```

```{r}
#depict barplot
p_ter_all_mol <- ggplot(data = ter_all_mol_longer, mapping = aes(x = Sample, y = emission, 
                                                         fill = factor(Terpene, 
                                                                       levels = names(ter_col_all)))) +
  geom_bar(stat = "identity", position = "fill") +
  facet_wrap(~Month, 
             scales = "free_x") +
  theme_classic(base_size = 11) + mytheme +
  scale_fill_manual(values = ter_col_all) +
  scale_y_continuous(labels = label_comma()) +
  labs(fill = "Terpene") +
  ylab("Relative basal emission rate") +
  theme(text = element_text(color = "black"), 
        strip.text = element_text(size = 6),
        legend.title = element_text(size = 8), 
        legend.text = element_markdown(size = 7),
        axis.title.x = element_blank(), 
        axis.text.x = element_text(size = 6, angle = 90))
show(p_ter_all_mol)

ggsave(plot = p_ter_all_mol, filename = paste0(path_out, "/figures/BVOC/BVOC_barrel_2007to2408_mol.pdf"), 
       dpi = 300, width = 160, height = 120, units = "mm")
```

```{r}
p_tertot_all_mol <- plot_bar(ps_ter_mol_pr, fill = "mol_weight") +
  geom_bar(aes(fill=mol_weight), stat="identity", position="stack") +
  scale_fill_viridis_d(labels = c("Diterpene", "Monoterpene", "Sesquiterpene")) +
  scale_y_continuous(labels = label_comma()) +
  theme_classic(base_size = 11) + mytheme +
  facet_wrap(~Month, 
             scales = "free_x") +
  labs(fill = "Terpene") +
  ylab(expression(paste("Basal emission rate (nmol ", {gdw^-1}, " ", {h^-1}, ")"))) +
  theme(text = element_text(color = "black"), 
        strip.text = element_text(size = 6),
        legend.title = element_text(size = 8), 
        legend.text = element_markdown(size = 7),
        axis.title.x = element_blank(), 
        axis.text.x = element_text(size = 6, angle = 90))

ggsave(plot = p_tertot_all_mol, filename = paste0(path_out, "/figures/BVOC/BVOC_barabs_2007to2408_mol.pdf"), 
       dpi = 300, width = 160, height = 120, units = "mm")
```

##Ohta
```{r}
#set the color of each terpene
ter_col_Ohta <- as.character(tax_table(ps_Ohta)[, "col"])
names(ter_col_Ohta) <- as.character(tax_table(ps_Ohta)[, "Terpene"])

#elong the dataset
ter_Ohta_mol_longer <- as.data.frame(otu_table(ps_Ohta_mol)) %>%
  mutate(Sample = rownames(.)) %>%
  pivot_longer(cols = -Sample, names_to = "Terpene", values_to = "emission")

#add metadata
meta_Ohta_mol <- data.frame(sample_data(ps_Ohta_mol)) %>%
  mutate(Sample = rownames(.))

ter_Ohta_mol_longer <-  ter_Ohta_mol_longer %>%
  left_join(meta_Ohta_mol, by = "Sample")
```

```{r}
#depict barplot
p_ter_mol <- ggplot(data = ter_Ohta_mol_longer, mapping = aes(x = Sample, y = emission, 
                                                       fill = factor(Terpene, 
                                                                     levels = names(ter_col_Ohta)))) +
  geom_bar(stat = "identity", position = "fill") +
  facet_wrap(~Month, 
             scales = "free_x") +
  theme_classic(base_size = 11) + mytheme +
  scale_fill_manual(values = ter_col_Ohta) +
  scale_y_continuous(labels = label_comma()) +
  labs(fill = "Terpene") +
  ylab("Relative basal emission rate") +
  theme(text = element_text(color = "black"), 
        strip.text = element_text(size = 6),
        legend.title = element_text(size = 8), 
        legend.text = element_markdown(size = 7),
        axis.title.x = element_blank(), 
        axis.text.x = element_text(size = 6, angle = 90))
show(p_ter)

ggsave(plot = p_ter_mol, filename = paste0(path_out, "/figures/BVOC/BVOC_barrel_2209to2408_mol.pdf"), 
       dpi = 300, width = 160, height = 120, units = "mm")
```

```{r}
###total basal emission rate
p_tertot_mol <- plot_bar(ps_Ohta_mol, fill = "mol_weight") +
  geom_bar(aes(fill=mol_weight), stat="identity", position="stack") +
  scale_fill_viridis_d(labels = c("Diterpene", "Monoterpene", "Sesquiterpene")) +
  scale_y_continuous(labels = label_comma()) +
  theme_classic(base_size = 11) + mytheme +
  facet_wrap(~Month, 
             scales = "free_x") +
  labs(fill = "Terpene") +
  ylab(expression(paste("Basal emission rate (nmol ", {gdw^-1}, " ", {h^-1}, ")"))) +
  theme(text = element_text(color = "black"), 
        strip.text = element_text(size = 6),
        legend.title = element_text(size = 8), 
        legend.text = element_markdown(size = 7),
        axis.title.x = element_blank(), 
        axis.text.x = element_text(size = 6, angle = 90))

ggsave(plot = p_tertot_mol, filename = paste0(path_out, "/figures/BVOC/BVOC_barabs_2209to2408_mol.pdf"), 
       dpi = 300, width = 160, height = 120, units = "mm")
```

#YET
#pca

```{r}
#remove kaurene from dataframe
#sapply(c("kaurene"), grep, colnames(BVOC_log_ps))
pca_BVOC_log_rm <- prcomp(BVOC_2306_log[, -c(14)], 
                          scale. = TRUE)
summary(pca_BVOC_log_rm)

#save(pca_BVOC_log_rm, file = paste0(path_out, "/231202_Sugi_phyllosphere_June_16S/processed_data/BVOC/pca_BVOC_2306_16S_log.Rdata"))
```

```{r}
col_BVOC <- c(rep("green3", 12), "purple3", "blue3")
col_pop <- colorRampPalette(c("dodgerblue", "orangered"))(3)

p_pca <- autoplot(pca_BVOC_log_rm, 
                  data = cbind(BVOC_2306_log[, -c(14)], 
                               population = metadata$population), 
                  x = 1, y = 2,
                  color = "population", 
                  scale = 1
                  )

#scale loading vectors
#see https://rdrr.io/cran/ggfortify/src/R/plotlib.R
loadings_sc <- data.frame(PC1 = pca_BVOC_log_rm$rotation[, 1], 
                          PC2 = pca_BVOC_log_rm$rotation[, 2], 
                          label = rownames(pca_BVOC_log_rm$rotation))
scaler <- min(max(abs(p_pca$data$PC1)) / max(abs(loadings_sc$PC1)),
              max(abs(p_pca$data$PC1)) / max(abs(loadings_sc$PC2)))

loadings_sc[, 1:2] <- loadings_sc[, 1:2] * scaler * 0.8
loadings_sc$label <- gg_italic(loadings_sc$label, "-")
#adjustment for the label positions
loadings_sc$xadj <- c(-0.07, -0.07, -0.07, -0.03, -0.07, -0.07, 0, -0.07, -0.07, -0.07, -0.07, 0, 0)
loadings_sc$yadj <- c(-0.007, 0.14, 0.06, 0.015, 0.085, 0.013, -0.015, -0.048, 0.115, 0.0375, -0.072, -0.015, 0.02)

p_pca <- p_pca + geom_segment(data = loadings_sc, 
                              mapping = aes(x = 0, xend = PC1, y = 0, yend = PC2), 
                              arrow = arrow(length = unit(2.0, "mm")), 
                              color = col_BVOC[-14], 
                              linewidth = 0.5) +
  geom_text(data = loadings_sc, 
            mapping = aes(x = PC1 + xadj, y = PC2 + yadj, label = label),
            size = 5.5 * .352777778,
            color = "black",
            parse = TRUE) +
  coord_fixed(ratio = 1) +
  theme_classic(base_size = 11) +
  labs(color = "Population") +
  scale_color_manual(values = col_pop, labels = c("Ajigazawa", "Azouji", "Yakushima")) +
  mytheme + theme(legend.position = "none")

show(p_pca)

ggsave(plot = p_pca, filename = paste0(path_out, "/231202_Sugi_phyllosphere_June_16S/figures/pca_BVOC_log.png"), 
       dpi = 300, width = 160, height = 160, units = "mm")
```

#merge plot

```{r}
p_pca_none <- p_pca + theme(legend.position = "bottom", 
                            legend.title = element_text(size = 7), 
                            legend.text = element_text(size = 7))
legend <- gtable_filter(ggplotGrob(p_pca_none), "guide-box")
p_pca_none <- p_pca + theme(legend.position = "none")

pl_BVOC <- arrangeGrob(p_BVOC, top = create_label("A"))
pl_pca <- arrangeGrob(p_pca_none, top = create_label("B"))

layout <- rbind(c(1, 2), c(1, 3))

p_merge <- grid.arrange(pl_BVOC, pl_pca, legend, 
                        layout_matrix = layout, 
                        widths = c(5.5, 4.5), 
                        heights = c(9, 1)
                        )

ggsave(plot = p_merge, filename = paste0(path_out, "/231202_Sugi_phyllosphere_June_16S/figures/barplot_BVOC_pca.png"), 
       dpi = 300, width = 160, height = 80, units = "mm", scale = 1)

#Don't do below command after editing with Inkscape
#ggsave(plot = p_merge, filename = paste0(path_out, #"/231202_Sugi_phyllosphere_June_16S/figures/barplot_BVOC_pca1.pdf"), 
#       dpi = 300, width = 160, height = 80, units = "mm", scale = 1, family = "Helvetica")
```

#LM (to test the correlations between latitude and BVOCs emission rates)

```{r}
#to test the differences of PC1, 2 among groups, perform ANOVA test
#merge metadata and the result of PCA
data_pca <- rownames_to_column(data.frame(pca_BVOC_log_rm$x))
data_pca <- merge(metadata, data_pca, by.x = "samplename", by.y = "rowname")

#merge log-transformed total basal emission rate
df_total <- data.frame(total = BVOC_2306_total_log)
df_total <- rownames_to_column(df_total)
data_pca <- merge(data_pca, df_total, by.x = "samplename", by.y = "rowname")

#merge BVOCs emission profile
df_BVOC <- rownames_to_column(BVOC_2306_log)
data_pca <- merge(data_pca, df_BVOC, by.x = "samplename", by.y = "rowname")

#merge population metadata
data_pca <- merge(data_pca, meta_pop, by.x = "population", by.y = "population", all.x = TRUE)
```

```{r}
#full model (use PC1 and PC2 for climatic data)
model_pc1_clim <- lm(PC1 ~ PC1_clim + PC2_clim, data = data_pca)
model_pc2_clim <- lm(PC2 ~ PC1_clim + PC2_clim, data = data_pca)
model_total_clim <- lm(total ~ PC1_clim + PC2_clim, data = data_pca)

anova(model_pc1_clim)
anova(model_pc2_clim)
anova(model_total_clim)
#none of the models are significant when both PC1 & PC2 included

write.csv(anova(model_pc1_clim), 
          file = paste0(path_out, "/231202_Sugi_phyllosphere_June_16S/analysis/BVOC_pc1_glm_climtot.csv"))
write.csv(anova(model_pc2_clim), 
          file = paste0(path_out, "/231202_Sugi_phyllosphere_June_16S/analysis/BVOC_pc2_glm_climtot.csv"))

write.csv(anova(model_total_clim), 
          file = paste0(path_out, "/231202_Sugi_phyllosphere_June_16S/analysis/BVOC_total_glm_climtot.csv"))
```

##test multicollinearity

```{r}
vif(model_pc1_clim)
#multicollinearity is not detected

#perform model selection and compare results with one covariate models
set.seed(123)
selc_PC1_clim <- stepAIC(model_pc1_clim)
selc_PC2_clim <- stepAIC(model_pc2_clim)
selc_total_clim <- stepAIC(model_total_clim)
#PC1 for climatic data was selected only for the terpene PC1 model
```

###climatic PC1

```{r}
#output models with climatic PC1 as an explanatory variable
model_pc1_clim <- lm(PC1 ~ PC1_clim, data = data_pca)
model_pc2_clim <- lm(PC2 ~ PC1_clim, data = data_pca)
model_total_clim <- lm(total ~ PC1_clim, data = data_pca)

write.csv(cbind(summary(model_pc1_clim)$coefficients, 
                AIC = summary(model_pc1_clim)$aic), 
          file = paste0(path_out, "/231202_Sugi_phyllosphere_June_16S/analysis/BVOC_pc1_glm_clim.csv"))

write.csv(cbind(summary(model_pc2_clim)$coefficients, 
                AIC = summary(model_pc2_clim)$aic), 
          file = paste0(path_out, "/231202_Sugi_phyllosphere_June_16S/analysis/BVOC_pc2_glm_clim.csv"))

write.csv(cbind(summary(model_total_clim)$coefficients, 
                AIC = summary(model_total_clim)$aic), 
          file = paste0(path_out, "/231202_Sugi_phyllosphere_June_16S/analysis/BVOC_total_glm_clim.csv"))

PC1clim <- rbind(summary(model_pc1_clim)$coefficient, 
                 summary(model_pc2_clim)$coefficient, 
                 summary(model_total_clim)$coefficient)
rownames(PC1clim) <- paste0(rep(c("PC1", "PC2", "total"), each = 2),
                            rep(c("Intercept", "PC1_clim"), 3)
                            )
```

###climatic PC2

```{r}
#output models with climatic PC2 as an explanatory variable
model_pc1_clim <- lm(PC1 ~ PC2_clim, data = data_pca)
model_pc2_clim <- lm(PC2 ~ PC2_clim, data = data_pca)
model_total_clim <- lm(total ~ PC2_clim, data = data_pca)

write.csv(cbind(summary(model_pc1_clim)$coefficients, 
                AIC = summary(model_pc1_clim)$aic), 
          file = paste0(path_out, "/231202_Sugi_phyllosphere_June_16S/analysis/BVOC_pc1_glm_clim2.csv"))

write.csv(cbind(summary(model_pc2_clim)$coefficients, 
                AIC = summary(model_pc2_clim)$aic), 
          file = paste0(path_out, "/231202_Sugi_phyllosphere_June_16S/analysis/BVOC_pc2_glm_clim2.csv"))

write.csv(cbind(summary(model_total_clim)$coefficients, 
                AIC = summary(model_total_clim)$aic), 
          file = paste0(path_out, "/231202_Sugi_phyllosphere_June_16S/analysis/BVOC_total_glm_clim2.csv"))

PC2clim <- rbind(summary(model_pc1_clim)$coefficient, 
                 summary(model_pc2_clim)$coefficient, 
                 summary(model_total_clim)$coefficient)
rownames(PC2clim) <- paste0(rep(c("PC1", "PC2", "total"), each = 2),
                            rep(c("Intercept", "PC2_clim"), 3)
                            )
```

##plot LM with climatic PCs

```{r}
#climatic PC1
p_pc1pc1 <- ggplot(data = data_pca, mapping = aes(x = PC1_clim, y = PC1, color = population)) +
  geom_smooth(method = "lm", color = "black", linewidth = .7, alpha = 0.1) +
  geom_point() +
  labs(x = "Climatic PC1", y = "Terpene PC1", color = "Population") +
  scale_color_manual(values = col_pop, labels = c("Ajigasawa", "Azouji", "Yakushima")) +
  theme_classic(base_size = 12) + mytheme +
  NULL

p_pc2pc1 <- ggplot(data = data_pca, mapping = aes(x = PC1_clim, y = PC2, color = population)) +
  geom_point() +
  labs(x = "Climatic PC1", y = "Terpene PC2", color = "Population") +
  scale_color_manual(values = col_pop, labels = c("Ajigasawa", "Azouji", "Yakushima")) +
  theme_classic(base_size = 12) + mytheme +
  NULL

#climatic PC2
p_pc1pc2 <- ggplot(data = data_pca, mapping = aes(x = PC2_clim, y = PC1, color = population)) +
  geom_smooth(method = "lm", color = "black", linewidth = .7, alpha = 0.1) +
  geom_point() +
  labs(x = "Climatic PC2", y = "Terpene PC1", color = "Population") +
  scale_color_manual(values = col_pop, labels = c("Ajigasawa", "Azouji", "Yakushima")) +
  theme_classic(base_size = 12) + mytheme +
  NULL

p_pc2pc2 <- ggplot(data = data_pca, mapping = aes(x = PC2_clim, y = PC2, color = population)) +
  geom_point() +
  labs(x = "Climatic PC2", y = "Terpene PC2", color = "Population") +
  scale_color_manual(values = col_pop, labels = c("Ajigasawa", "Azouji", "Yakushima")) +
  theme_classic(base_size = 12) + mytheme +
  NULL

show(p_pc1pc1)
show(p_pc2pc1)
show(p_pc1pc2)
show(p_pc2pc2)
```

```{r}
##arrange plots
pl_pc2pc1 <- p_pc2pc1 + theme(legend.position = "bottom")
legend <- gtable_filter(ggplotGrob(pl_pc2pc1), "guide-box")
pl_pc1pc1 <- p_pc1pc1 + theme(legend.position = "none")
pl_pc2pc1 <- p_pc2pc1 + theme(legend.position = "none")
pl_pc1pc2 <- p_pc1pc2 + theme(legend.position = "none")
pl_pc2pc2 <- p_pc2pc2 + theme(legend.position = "none")

pl_pc1pc1 <- arrangeGrob(pl_pc1pc1, top = create_label("A"))
pl_pc1pc2 <- arrangeGrob(pl_pc1pc2, top = create_label("B"))
pl_pc2pc1 <- arrangeGrob(pl_pc2pc1, top = create_label("C"))
pl_pc2pc2 <- arrangeGrob(pl_pc2pc2, top = create_label("D"))

layout <- rbind(c(1, 2), c(3, 4), c(5, 5))

p_lm <- grid.arrange(pl_pc1pc1, pl_pc1pc2, pl_pc2pc1, pl_pc2pc2, legend,
                     layout_matrix = layout, widths = c(1, 1), heights = c(9, 9, 2))

ggsave(plot = p_lm, filename = paste0(path_out, "/231202_Sugi_phyllosphere_June_16S/figures/LM_terpene_climate1.pdf"), 
       dpi = 300, width = 160, height = 120, units = "mm", scale = 1)
```

##adjusted p value

```{r}
adjust <- rbind(PC1clim[c(2, 4), ], PC2clim[c(2, 4), ])
adjust <- cbind(adjust, p_adj = p.adjust(adjust[, 4], method = "fdr"))
```

```{r}
model_pc1 <- glm(PC1 ~ latitude, data = data_pca)
model_pc2 <- glm(PC2 ~ latitude, data = data_pca)
model_total <- glm(total ~ latitude, data = data_pca)

summary(model_pc1)
summary(model_pc2)
summary(model_total)

write.csv(cbind(summary(model_pc1)$coefficients, 
                AIC = summary(model_pc1)$aic), 
          file = paste0(path_out, "/231202_Sugi_phyllosphere_June_16S/analysis/BVOC_pc1_glm.csv"))

write.csv(cbind(summary(model_pc2)$coefficients, 
                AIC = summary(model_pc2)$aic), 
          file = paste0(path_out, "/231202_Sugi_phyllosphere_June_16S/analysis/BVOC_pc2_glm.csv"))

write.csv(cbind(summary(model_total)$coefficients, 
                AIC = summary(model_total)$aic), 
          file = paste0(path_out, "/231202_Sugi_phyllosphere_June_16S/analysis/BVOC_total_glm.csv"))
```

