---
title: "BVOC_ILRandPCA_gram"
output: html_document
date: "2025-02-27"
---

Perform log-ratio transformation and PCA on gram-based data

#load

```{r}
library(readxl)
library(phyloseq)
library(tidyverse)
library(ggfortify)
library(RColorBrewer)
```

#path to raw-data directory (change "path" to your raw-data directory)

```{r}
#create "path" file, and define "path" as your raw-data directory(character vector) in the file
#"path" is raw-data directory, and "path_out" is output file directory

source("../scripts/path.R")
```

#source codes

```{r}
#source("../R/label.R")
#source("../R/greek_as_alphabet.R")
```

#theme settings

```{r}
#theme settings
mytheme <-theme(text = element_text(color = "black"), 
                axis.text = element_text(color = "black"), 
                legend.key.size = unit(8, "pt"), 
                legend.spacing.x = unit(0, "mm"))
```

#import data

```{r}
load(file = paste0(path_out, "/data/Sugi/BVOC/phyloseq_Terpene_2007to2408_imp.Rdata"))
load(file = paste0(path_out, "/data/Sugi/BVOC/Terpene_stored_K2007.Rdata"))

#extract Yoshioka data

ps_ter_imp3p <- tax_glom(ps_ter_imp, taxrank = "mol_weight")
taxa_names(ps_ter_imp3p) <- c("Monoterpene", "Sesquiterpene", "Diterpene")
```

##merge heavy Terpene

```{r}
#The emitted sesquiterpenes and diterpenes are dominated by beta-Farnesene and ent-Kaurene
#Thus, merge Sesquiterpene and Diterpene
ps_ter_mergeHeavy <- merge_taxa(ps_ter_imp,
                                taxa_names(ps_ter_imp)[tax_table(ps_ter_imp)[, "mol_weight"] == "Sesqui"],
                                "beta-Farnesene")
tax_table(ps_ter_mergeHeavy)["beta-Farnesene", "col"] <- tax_table(ps_ter_imp)["beta-Farnesene", "col"]
ps_ter_mergeHeavy <- merge_taxa(ps_ter_mergeHeavy,
                                taxa_names(ps_ter_mergeHeavy)[tax_table(ps_ter_mergeHeavy)[, "mol_weight"] == "Di"],
                                "ent-Kaurene")
tax_table(ps_ter_mergeHeavy)["ent-Kaurene", "col"] <- tax_table(ps_ter_imp)["ent-Kaurene", "col"]
```

#log-transform

```{r}
ps_ter_log <- transform_sample_counts(ps_ter_imp, log)
ps_ter_log_mergeHeavy <- transform_sample_counts(ps_ter_mergeHeavy, log)

#3 parts
ps_ter_log3p <- transform_sample_counts(ps_ter_imp3p, log)
```

###Yoshioka data

```{r}
ps_Yoshioka_log <- subset_samples(ps_ter_log, Month == "2007")
ps_Yoshioka_log_mergeHeavy <- subset_samples(ps_ter_log_mergeHeavy, Month == "2007")
ps_Yoshioka_log3p <- subset_samples(ps_ter_log3p, Month == "2007")
```

#scatter plot
##3 parts

```{r}
pldata_Yoshi3p <- data.frame(otu_table(ps_Yoshioka_log3p))
pldata_Yoshi3p <- rownames_to_column(pldata_Yoshi3p, var = "sampleID")
pldata_Yoshi3p <- left_join(pldata_Yoshi3p, data.frame(sample_data(ps_Yoshioka_log3p)), by = "sampleID")

p_Yoshi_MS <- ggplot(data = pldata_Yoshi3p, aes(x = Monoterpene, y = Sesquiterpene, 
                                            color = PC1_clim)) + 
  geom_point(size = 3) +
  theme_bw(base_size = 11) + mytheme +
  scale_color_viridis_c()

show(p_Yoshi_MS)

#ggsave(p_Yoshi_MS, 
#       file = paste0(path_out, "231202_Sugi_phyllosphere_June_16S/figures/BVOC/BVOC_log_2007_MS.pdf"), 
#       width = 160, height = 120, unit = "mm", dpi = 300)
```

```{r}
p_Yoshi_MD <- ggplot(data = pldata_Yoshi3p, aes(x = Monoterpene, y = Diterpene, 
                                            color = PC1_clim)) + 
  geom_point(size = 3) +
  theme_bw(base_size = 11) + mytheme +
  scale_color_viridis_c()

show(p_Yoshi_MD)
#ggsave(p_Yoshi_MD, 
#       file = paste0(path_out, "231202_Sugi_phyllosphere_June_16S/figures/BVOC/BVOC_log_2007_MD.pdf"), 
#       width = 160, height = 120, unit = "mm", dpi = 300)
```

##write transformed

```{r}
save(ps_Yoshioka_log_mergeHeavy, 
     file = paste0(path_out, "/data/Sugi/BVOC/phyloseq_2007_log_mergeHeavy.Rdata"))
save(ps_Yoshioka_log, 
     file = paste0(path_out, "/data/Sugi/BVOC/phyloseq_2007_log.Rdata"))
save(ps_Yoshioka_log3p, 
     file = paste0(path_out, "/data/Sugi/BVOC/phyloseq_2007_log3p.Rdata"))
```

##PCA (merged)

```{r}
#extract data
log_Yoshioka_mergeHeavy <- as.matrix(otu_table(ps_Yoshioka_log_mergeHeavy))
log_Yoshioka <- as.matrix(otu_table(ps_Yoshioka_log))
```

```{r}
pca_log_Yoshioka <- prcomp(log_Yoshioka, scale. = TRUE, center = TRUE)
summary(pca_log_Yoshioka)
pca_log_Yoshioka_mergeHeavy <- prcomp(log_Yoshioka_mergeHeavy, scale. = TRUE, center = TRUE)
summary(pca_log_Yoshioka_mergeHeavy)
plot(pca_log_Yoshioka)
plot(pca_log_Yoshioka_mergeHeavy)

save(pca_log_Yoshioka_mergeHeavy, 
     file = paste0(path_out, "/data/Sugi/BVOC/pcalog_2007_mergeHeavy.Rdata"))
save(pca_log_Yoshioka, 
     file = paste0(path_out, "/data/Sugi/BVOC/pcalog_2007.Rdata"))
```

###plot

```{r}
#plot PC1, 2
p_logpca12_Yoshioka_merHev <- autoplot(pca_log_Yoshioka_mergeHeavy,
                                       data = data.frame(sample_data(ps_Yoshioka_log_mergeHeavy)),
                                       color = "PC1_clim",
                                       loadings = TRUE,
                                       loadings.color = tax_table(ps_Yoshioka_log_mergeHeavy)[, "col"],
                                       loadings.label = TRUE,
                                       loadings.label.color = "black",
                                       loadings.label.repel=TRUE) +
  scale_color_viridis_c() +
  coord_fixed(1) +
  #xlim(-0.3, 0.3) + ylim(-0.3, 0.3) +
  theme_classic(base_size = 12) + mytheme

show(p_logpca12_Yoshioka_merHev)
```


**The heavy terpenes are negatively correlated with the major MTs, as the Ota data.**

```{r}
#plot PC2, 3
p_logpca23_Yoshioka_merHev <- autoplot(pca_log_Yoshioka_mergeHeavy,
                                       x = 2, y = 3,
                                       loadings = TRUE,
                                       data = data.frame(sample_data(ps_Yoshioka_log_mergeHeavy)),
                                       color = "PC1_clim",
                                       loadings.color = tax_table(ps_Yoshioka_log_mergeHeavy)[, "col"],
                                       loadings.label = TRUE,
                                       loadings.label.color = "black",
                                       loadings.label.repel=TRUE) +
  scale_color_viridis_c() +
  coord_fixed(1) +
  theme_classic(base_size = 12) + mytheme

show(p_logpca23_Yoshioka_merHev)
```

```{r}
show(p_logpca12_Yoshioka_merHev)
show(p_logpca23_Yoshioka_merHev)
##PC4 is affected by some outliers

##save plots
ggsave(p_logpca12_Yoshioka_merHev, 
       file = paste0(path_out, "/231202_Sugi_phyllosphere_June_16S/figures/BVOC_logpca12_2007.pdf"), 
       width = 160, height = 160, unit = "mm", dpi = 300)
ggsave(p_logpca23_Yoshioka_merHev, 
       file = paste0(path_out, "/231202_Sugi_phyllosphere_June_16S/figures/BVOC_logpca23_2007.pdf"), 
       width = 160, height = 160, unit = "mm", dpi = 300)
```

**The overall trend is similar to that of Ota data. MTs(a-Pinene, Limonene) and SQTs&DTs are negatively correlated.**
