---
title: "FungalTraits"
output: html_document
date: "2025-07-07"
---
#load packages

```{r}
#test commit
library(readxl)
library(phyloseq)
library(tidyverse)
library(vegan)
library(glm2)
library(MASS)
library(lmtest)
library(ggfortify)
library(ggeffects)
library(RColorBrewer)
library(gridExtra)
library(grid)
library(gtable)
library(gt)
```

#path to raw-data directory (change "path" to your raw-data directory)

```{r}
#create "path" file, and define "path" as your raw-data directory(character vector) in the file
#"path" is raw-data directory, and "path_out" is output file directory

source("../scripts/path.R")
```

#source codes

```{r}
source("../R/occurrence.R")
source("../R/label.R")
source("../R/Greek.R")
source("../R/html_italic.R")
```

##mytheme

```{r}
#theme settings
mytheme <-theme(text = element_text(color = "black"), 
                axis.text = element_text(color = "black"),
                axis.title.y = element_text(size = 8), 
                
                legend.key.size = unit(9, "pt"), 
                legend.spacing.x = unit(0, "mm"))
```

#Guild based on the Kobayashi data

```{r}
path_gs <- read.csv(paste0(path_out, "/231202_Sugi_phyllosphere_June_ITS/analysis/pathogen_Gs.csv"))
path_gs <- column_to_rownames(path_gs, var = "X")
FUNGuild_path <- read_tsv(paste0(path_out, "/231202_Sugi_phyllosphere_June_ITS/processed_data/FUNGuild/otu_table_Kobayashi.guilds.txt"))
path_gs_out <- read_tsv(file = paste0(path_out, "/231202_Sugi_phyllosphere_June_ITS/export/Pathogen_putative.tsv"))

load(paste0(path_out, "/231202_Sugi_phyllosphere_June_ITS/processed_data/phyloseq_pathogen.Rdata"))
```

**Important: the database "FungalTraits" is different from "fungal_traits" (also called "fun<sup>fun</sup>").**
\n**Use "FungalTraitss"**

```{r}
#FungalTraits assumes the ITS sequencing data with species-level taxonomic information for guild assignment, but the targeted ASVs are assigned to the genus level taxonomies only.
FunTr <- read.csv(paste0(path_out, "/data/Sugi/FungalTraits/FungalTraits 1.2_ver_16Dec_2020 - V.1.2.csv"))

FunTr_path <- FunTr %>%
  dplyr::filter(GENUS %in% path_gs_out$Genus) %>%
  dplyr::select(GENUS, 
                primary_lifestyle, 
                Secondary_lifestyle, 
                Comment_on_lifestyle_template, 
                Plant_pathogenic_capacity_template, 
                Decay_substrate_template, 
                Decay_type_template)

path_gs_both <- dplyr::left_join(path_gs_out, 
                                 FunTr_path, 
                                 by = c("Genus" = "GENUS"))
show(path_gs_both)
```

**All of the fungal genera listed in Kobayashi 2007 included plant saprotrophs other than <i>Tremella</i>.**
#GLM
##Saprotroph

```{r}
#subset saprotrophs
ps_sapro <- subset_taxa(ps_ITS_path, Genus != "Tremella")

#aggregate saprotrophic ASVs
sapro <- rownames_to_column(data.frame(Saprotroph = sample_sums(ps_sapro)), var = "ID")
meta_sapro <- data.frame(sample_data(ps_sapro)) %>%
  dplyr::select(c(site, latitude, PC1_clim, PC2_clim)) %>%
  rownames_to_column(var = "ID")

pldata_sapro <- left_join(sapro, meta_sapro, by = "ID") %>%
  column_to_rownames(var = "ID")
```

```{r}
#GLM for saprotrophs
model_sapro <- glm.nb(Saprotroph ~ site + PC1_clim, 
                      data = pldata_sapro, 
                      method = "glm.fit", 
                      link = log)

summary(model_sapro)
```

###plot

```{r}
#show prediction and CI for the model with the total abundance tested on PC1_clim
model_pred <- glm.nb(Saprotroph ~ PC1_clim, 
                      data = pldata_sapro, 
                      method = "glm.fit", 
                      link = log)
new_data <- data.frame(PC1_clim = seq(min(pldata_sapro$PC1_clim), 
                                      max(pldata_sapro$PC1_clim), 
                                      length.out = 100))
pred <- predict(model_pred, newdata = new_data, type = "link", se.fit = TRUE)
crit <- qnorm(0.975)  # For 95% confidence intervals

new_data$fit <- exp(pred$fit)
new_data$lower <- exp(pred$fit - crit * pred$se.fit)
new_data$upper <- exp(pred$fit + crit * pred$se.fit)

p_sapro <- ggplot(data = pldata_sapro, 
                 mapping = aes(x = PC1_clim, y = Saprotroph)) + 
  #scale_color_manual(values = guild$col[sapply(levels(abund_ga$FuncType), grep, guild$FuncType)]) + 
  geom_line(data = new_data, aes(y = fit), linewidth = 0.7, color = "black") +
  geom_ribbon(data = new_data, 
              aes(ymin = lower, ymax = upper, y = fit),
              alpha = 0.05, linewidth = 0, color = NA, fill = "black") +
  geom_point() + 
  theme_classic() + mytheme +
  xlab("Climatic PC1") + ylab("Abundance of the putative saprotrophic fungi after rarefaction") + 
  #labs(color = "Putative functional group") + 
  NULL
  
show(p_sapro)

ggsave(filename = paste0(path_out, "/231202_Sugi_phyllosphere_June_ITS/figures/sapro_climate1.png"), 
       plot = p_sapro, 
       dpi = 300, width = 160, height = 90, units = "mm")
```

##Pathogen

```{r}
#subset the genera which include plant pathogenic fungi
path_path <- path_gs_both %>%
  dplyr::filter(grepl("pathogen", primary_lifestyle) |
                  grepl("pathogen", Secondary_lifestyle) | 
                  grepl("pathogen", Comment_on_lifestyle_template) | 
                  grepl("pathogen", Plant_pathogenic_capacity_template))

#subset pathogen ASVs
ps_path <- subset_taxa(ps_ITS_path, Genus %in% path_path$Genus)

#aggregate saprotrophic ASVs
pathogen <- rownames_to_column(data.frame(Pathogen = sample_sums(ps_path)), var = "ID")
meta_path <- data.frame(sample_data(ps_path)) %>%
  dplyr::select(c(site, latitude, PC1_clim, PC2_clim)) %>%
  rownames_to_column(var = "ID")

pldata_path <- left_join(pathogen, meta_path, by = "ID") %>%
  column_to_rownames(var = "ID")
```

```{r}
#GLM for saprotrophs
model_path <- glm.nb(Pathogen ~ site + PC1_clim, 
                      data = pldata_path, 
                      method = "glm.fit", 
                      link = log)

summary(model_path)
```

###plot

```{r}
#show prediction and CI for the model with the total abundance tested on PC1_clim
model_pred <- glm.nb(Pathogen ~ PC1_clim, 
                      data = pldata_path, 
                      method = "glm.fit", 
                      link = log)
new_data <- data.frame(PC1_clim = seq(min(pldata_path$PC1_clim), 
                                      max(pldata_path$PC1_clim), 
                                      length.out = 100))
pred <- predict(model_pred, newdata = new_data, type = "link", se.fit = TRUE)
crit <- qnorm(0.975)  # For 95% confidence intervals

new_data$fit <- exp(pred$fit)
new_data$lower <- exp(pred$fit - crit * pred$se.fit)
new_data$upper <- exp(pred$fit + crit * pred$se.fit)

p_path <- ggplot(data = pldata_path, 
                 mapping = aes(x = PC1_clim, y = Pathogen)) + 
  #scale_color_manual(values = guild$col[sapply(levels(abund_ga$FuncType), grep, guild$FuncType)]) + 
  geom_line(data = new_data, aes(y = fit), linewidth = 0.7, color = "black") +
  geom_ribbon(data = new_data, 
              aes(ymin = lower, ymax = upper, y = fit),
              alpha = 0.05, linewidth = 0, color = NA, fill = "black") +
  geom_point() + 
  theme_classic() + mytheme +
  xlab("Climatic PC1") + ylab("Abundance of the putative pathogenic fungi after rarefaction") + 
  #labs(color = "Putative functional group") + 
  NULL
  
show(p_path)

ggsave(filename = paste0(path_out, "/231202_Sugi_phyllosphere_June_ITS/figures/path_climate1.png"), 
       plot = p_path, 
       dpi = 300, width = 160, height = 90, units = "mm")
```

##res_GLM

```{r}
summary(model_path)
summary(model_sapro)
write.csv(cbind(call = paste(summary(model_path)$call, collapse = ", "), 
                coefficinets = summary(model_path)$coefficients, 
                aic = summary(model_path)$aic), 
          paste0(path_out, "/231202_Sugi_phyllosphere_June_ITS/analysis/glm_plant-pathogen.csv"))
write.csv(cbind(call = paste(summary(model_sapro)$call, collapse = ", "), 
                coefficinets = summary(model_sapro)$coefficients, 
                aic = summary(model_sapro)$aic), 
          paste0(path_out, "/231202_Sugi_phyllosphere_June_ITS/analysis/glm_sapro.csv"))
```

**The results suggested negative correlations between the abundances of both functional groups and PC1_clim.**
\n**This is because the dominant fungal genera (<i>Trametes</i> and <i>Diaporthe</i>) include both plant pathogenic and saprotrophic species.**

##plot_both

```{r}
#show the abundances of both putative saprotrophic and pathogenic fungi in the same plot
pldata_both <- left_join(sapro, pathogen, by = "ID") %>%
  left_join(meta_sapro, by = "ID") %>%
  column_to_rownames(var = "ID")

pldata_both <- pldata_both %>%
  pivot_longer(cols = c("Saprotroph", "Pathogen"), 
                        names_to = "FuncType", 
                        values_to = "Abundance")
```

```{r}
#show prediction and CI for the model with the total abundance tested on PC1_clim
model_pred <- glm.nb(Abundance ~ FuncType + PC1_clim, 
                      data = pldata_both, 
                      method = "glm.fit", 
                      link = log)
new_data <- data.frame(PC1_clim = rep(seq(min(pldata_path$PC1_clim), 
                                      max(pldata_path$PC1_clim), 
                                      length.out = 100), 2), 
                       FuncType = rep(c("Saprotroph", "Pathogen"), each = 100))
pred <- predict(model_pred, newdata = new_data, type = "link", se.fit = TRUE)
crit <- qnorm(0.975)  # For 95% confidence intervals

new_data$fit <- exp(pred$fit)
new_data$lower <- exp(pred$fit - crit * pred$se.fit)
new_data$upper <- exp(pred$fit + crit * pred$se.fit)

p_both <- ggplot(data = pldata_both, 
                 mapping = aes(x = PC1_clim, y = Abundance, colour = FuncType, shape = FuncType)) + 
  scale_color_manual(values = c(Saprotroph = "blue3", Pathogen = "rosybrown2")) + 
  geom_line(data = new_data, aes(y = fit, colour = FuncType), linewidth = 0.7) +
  geom_ribbon(data = new_data, 
              aes(ymin = lower, ymax = upper, y = fit, fill = FuncType),
              alpha = 0.05, linewidth = 0, color = NA) +
  geom_point() + 
  theme_classic() + mytheme +
  xlab("Climatic PC1") + ylab("Abundance of the putative Japanese-cedar-associated fungi \nafter rarefaction") + 
  labs(color = "Putative \nfunctional group", 
       fill = "Putative \nfunctional group", 
       shape = "Putative \nfunctional group") + 
  NULL
  
show(p_both)

ggsave(filename = paste0(path_out, "/231202_Sugi_phyllosphere_June_ITS/figures/sapro-path_climate1.png"), 
       plot = p_both, 
       dpi = 300, width = 160, height = 90, units = "mm")
```

#YET TO BE DONE*
**Output a table to show the result of the functional assignment using the FungalTraits database.**
