---
title: "pathogen_Hiura2021"
output: html_document
date: "2024-10-10"
---

#load packages

```{r}
library(readxl)
library(phyloseq)
library(tidyverse)
library(vegan)
library(glm2)
library(MASS)
library(RColorBrewer)
library(gridExtra)
library(grid)
library(gtable)
library(gt)
```

#path to raw-data directory (change "path" to your raw-data directory)

```{r}
#create "path" file, and define "path" as your raw-data directory(character vector) in the file
#"path" is raw-data directory, and "path_out" is output file directory

source("../scripts/path.R")
```

#source codes

```{r}
source("../R/occurrence.R")
source("../R/label.R")
```

#read processed data
#ITS

```{r import}
#rarefied_ASV_table
df_ITS <- read.csv(paste0(path_out, "/231202_Sugi_phyllosphere_June_ITS/processed_data/df_rarefied_ITS.csv"), row.names = 1)
rownames(df_ITS) <- formatC(as.numeric(rownames(df_ITS)), width = 4, flag = "0") #add 0 to rownames

#tax_table
tax_ITS <- read.csv(paste0(path_out, "/231202_Sugi_phyllosphere_June_ITS/processed_data/tax_table2_ITS.csv"), row.names = 1)
tax_ITS <- as.matrix(tax_ITS)

#phylogenetic tree
load(paste0(path_out, "/231202_Sugi_phyllosphere_June_ITS/processed_data/trees.Rdata"))
treeITS <- tree2
treeITS_rarefied <- tree_rarefied2
rm(list = c("tree2", "tree_rarefied2"))

#metadata
metadata <- read_excel(paste0(path_out, "/data/Sugi/metadata_2306.xlsx"), sheet = "Sheet1")
metadata <- column_to_rownames(metadata, var = "Sample_Name")
metadata$cloneID <- as.factor(metadata$cloneID)
metadata$population <- as.factor(metadata$population)
metadata$site <- as.factor(metadata$site)
metadata$date <- as.factor(metadata$date)

#read climate data
meta_pop <- read.csv(paste0(path_out, "/data/Sugi/pop_climate_pca.csv"))
meta_pop <- column_to_rownames(meta_pop, var = "X")

#add climatic data
metadata <- metadata %>%
  dplyr::select(-c(fullname, latitude, longitude, altitude)) %>%
  left_join(meta_pop, by = "population")
rownames(metadata) <- formatC(as.numeric(rownames(metadata)), width = 4, flag = "0")

#read pathogen data (Hiura et al. 2021)
pathogen <- read.csv(file = paste0(path_out, "/231202_Sugi_phyllosphere_June_ITS/processed_data/115fungal_spp_guild.csv"))
pathogen$Genus <- sapply(str_split(pathogen$species, " "), `[`, 1)
pathogen$species <- str_replace(pathogen$species, " ", "_")

#load result of nmds
#load nmds output
load(paste0(path_out, "/231202_Sugi_phyllosphere_June_ITS/processed_data/NMDS_bray_ITS.Rdata"))
nmds_ITS <- nmds_bray
rm(nmds_bray)
```

##for BVOCs

```{r}
#the microbiome outlier (AZ_040_1, based on bray) is excluded in advance
#rarefied-ASV table
df_ITS_ter <- read.csv(paste0(path_out, "/231202_Sugi_phyllosphere_June_ITS/processed_data/BVOC/df_rarefied_ITS_BVOC.csv"), row.names = 1)
rownames(df_ITS_ter) <- formatC(as.numeric(rownames(df_ITS_ter)), width = 4, flag = "0") 
#add 0 to the head of rownames

#metadata
meta_ter <- read.csv(paste0(path_out, "/231202_Sugi_phyllosphere_June_ITS/processed_data/BVOC/metadata_ITS_BVOC.csv"), row.names = 1)
rownames(meta_ter) <- formatC(as.numeric(rownames(meta_ter)), width = 4, flag = "0")

meta_ter$cloneID <- as.factor(meta_ter$cloneID)
meta_ter$population <- as.factor(meta_ter$population)
meta_ter$site <- as.factor(meta_ter$site)
meta_ter$date <- as.factor(meta_ter$date)

#log-transformed BVOCs Emission
load(paste0(path_out, "/231202_Sugi_phyllosphere_June_ITS/processed_data/BVOC/BVOC_2306_ITS_log.Rdata"))

#output of NMDS (based on bray-curtis dissimilarity)
load(paste0(path_out, "/231202_Sugi_phyllosphere_June_ITS/processed_data/BVOC/NMDS_bray_ITS_BVOC2306.Rdata"))
nmds_ITSter <- nmds_bray
rm(nmds_bray)

#output of vector-fitting (BVOC on NMDS based on bray)
load(paste0(path_out, "/231202_Sugi_phyllosphere_June_ITS/analysis/envfit_bray_BVOC.Rdata"))
```

##FUNGuild

```{r}
FUNGuild_path <- read_tsv(paste0(path_out, "/231202_Sugi_phyllosphere_June_ITS/processed_data/FUNGuild/otu_table_Kobayashi.guilds.txt"))
```

#ps

```{r}
ps_ITS <- phyloseq(otu_table(df_ITS, taxa_are_rows = FALSE), 
               tax_table(tax_ITS), 
               phy_tree(treeITS_rarefied$data), 
               sample_data(metadata))
ps_ITS_ter <- phyloseq(otu_table(df_ITS_ter, taxa_are_rows = FALSE), 
                       tax_table(tax_ITS), 
                       phy_tree(treeITS_rarefied$data), 
                       sample_data(meta_ter))

#agglomerate at Species level
ps_ITS_sp <- tax_glom(ps_ITS, taxrank = "Species")
ps_ITS_ter_sp <- tax_glom(ps_ITS_ter, taxrank = "Species")

#agglomerate at Genus level
ps_ITS_gs <- tax_glom(ps_ITS, taxrank = "Genus")
ps_ITS_ter_gs <- tax_glom(ps_ITS_ter, taxrank = "Genus")
```

#search Species from tax table

```{r}
#search pathogens in phyllosphere microbiome
ps_ITS_sp_path <- subset_taxa(ps_ITS_sp, Species %in% pathogen$species)

path_sp <- pathogen[pathogen$species %in% tax_table(ps_ITS_sp_path)[, "Species"], ]
path_sp
#only one species "Sarea_resinae" was detected, whose function is unknown

occurrence(otu_table(ps_ITS_sp_path))
sample_data(subset_samples(ps_ITS_sp_path, (otu_table(ps_ITS_sp_path))[, 1] > 0))
#detected in IS_009 and SM_003 at Kawatabi

path_sp$occ <- paste0("IS_009_1", "-", "SM_003_1", "-Kawatabi")

#write.csv(path_sp, file = paste0(path_out, "/231202_Sugi_phyllosphere_June_ITS/analysis/pathogen_Sp.csv"))
```

##search Genus

```{r}
#summarize pathogen data
#some Genus contains two species with different function
path_gs1 <- pathogen %>%
  group_by(Genus) %>%
  summarize(Genus = first(Genus), 
            species = first(species), 
            FuncType = first(FuncType))
path_gs2 <- pathogen %>%
  group_by(Genus) %>%
  summarize(Genus = nth(Genus, 2), 
            species = nth(species, 2), 
            FuncType = nth(FuncType, 2))
path_gs <- cbind(path_gs1, 
                 species2 = path_gs2$species, FuncType2 = path_gs2$FuncType)
path_gs$FuncType2[is.na(path_gs$FuncType2)] <- path_gs$FuncType[is.na(path_gs$FuncType2)]

##ASVs classified as the Genera included in pathogen data
#test <- subset_taxa(ps_ITS, Genus %in% path_gs$Genus)

#path_gs3 <- pathogen %>%
#  group_by(Genus) %>%
#  summarize(Genus = nth(Genus, 3), 
#            species = nth(species, 3), 
#            FuncType = nth(FuncType, 3))
```

```{r}
ps_ITS_path <- subset_taxa(ps_ITS_gs, Genus %in% path_gs$Genus)

tax_table(ps_ITS_path)
#10 putative pathogenic Genera were also detected in the microbiome data
#extract these ASVs
path_gs <- path_gs %>%
  left_join(dplyr::select(rownames_to_column(data.frame(tax_table(ps_ITS_path)), 
                                             var = "ASV"), 
                          c("Genus", "ASV")), 
            by = "Genus")

#ps object with only Genera which include pathogens
ps_ITS_ter_path <- prune_taxa(taxa_names(ps_ITS_path), ps_ITS_ter_gs)

#occurrence of each Genus
path_gs <- path_gs %>%
  left_join(rownames_to_column(data.frame(occurrence = taxa_sums(otu_table(ps_ITS_path) > 0)), 
                               var = "ASV"), by = "ASV")

#write.csv(path_gs, file = paste0(path_out, "/231202_Sugi_phyllosphere_June_ITS/analysis/pathogen_Gs.csv"))
```


#agglomerate genera according to functional guilds

```{r}
#ps object in which taxa are agglomerated according to functional guilds
ps_ITS_guild <- ps_ITS_path
ps_ITS_guild2 <- ps_ITS_guild

#rows of each ASVs in path_gs
rows <- sapply(rownames(ps_ITS_guild@tax_table), grep, path_gs$ASV)

ps_ITS_guild@tax_table <- tax_table(cbind(FuncType = path_gs[rows, "FuncType"], 
                                          FuncType2 = path_gs[rows, "FuncType2"], 
                                          ps_ITS_path@tax_table))
ps_ITS_guild2@tax_table <- tax_table(cbind(FuncType2 = path_gs[rows, "FuncType2"], 
                                           FuncType = path_gs[rows, "FuncType"], 
                                           ps_ITS_path@tax_table))


#treat Xeromphalina as a root rot fungi
ps_ITS_guild1 <- tax_glom(ps_ITS_guild, taxrank = "FuncType")
#treat Xeromphalina as a brown rot fungi
ps_ITS_guild2 <- tax_glom(ps_ITS_guild2, taxrank = "FuncType2")

save(ps_ITS_path, ps_ITS_guild, ps_ITS_guild1, ps_ITS_guild2, file = paste0(path_out, "/231202_Sugi_phyllosphere_June_ITS/processed_data/phyloseq_pathogen.Rdata"))
```

##output pathogen table

```{r}
path_gs_out <- path_gs
path_gs_out$reference <- pathogen[sapply(path_gs$species, charmatch, pathogen$species), 7]
path_gs_out <- left_join(path_gs_out, FUNGuild_path[, c(1, 110:ncol(FUNGuild_path))], by = "ASV")
path_gs_out <- path_gs_out %>%
  tidyr::drop_na(ASV) %>%
  dplyr::select("Genus" = Genus,
                "Species" = species,
                "Functional group" = FuncType,
                "Reference" = reference, 
                "Trait" = Trait,
                "Guild" = Guild,
                "Trophic mode" = `Trophic Mode`,
                "Confidence" = `Confidence Ranking`,
                "Notes" = Notes)
path_gs_out$Species <- gsub("_", " ", path_gs_out$Species)

write_tsv(path_gs_out, file = paste0(path_out, "/231202_Sugi_phyllosphere_June_ITS/export/Pathogen_putative.tsv"))
```

```{r}
path_gs_tbl <- path_gs_out %>% 
  dplyr::select(-Notes) %>%
  #change the font for journal titles to italic
  dplyr::mutate(Reference = str_replace(Reference, "Mycologia 87", "<i>Mycologia</i> <b>87</b>") %>%
                  purrr::map(gt::md), 
                .keep = "unused"
                ) %>%
  dplyr::mutate(Reference = str_replace(Reference, "J. Wood Sci. 53", "<i>J. Wood Sci.</i> <b>53</b>") %>%
                  purrr::map(gt::md), 
                .keep = "unused"
                ) %>%
  dplyr::mutate(`Trophic mode` = str_replace(`Trophic mode`, "Pathotroph-Saprotroph-Symbiotroph", 
                                             "Pathotroph-Saprotroph-<br>Symbiotroph") %>%
                  purrr::map(gt::md), 
                .keep = "unused"
                ) %>%
  dplyr::mutate(Guild = str_replace(Guild, 
                                    "Plant Pathogen-Plant Saprotroph-Undefined Saprotroph-", 
                                    "Plant Pathogen-Plant Saprotroph-<br>Undefined Saprotroph-") %>%
                  purrr::map(gt::md), 
                .keep = "unused"
                ) %>%
  
  gt::gt() %>%
  gt::tab_style(cell_text(weight = "bold"), 
                locations = cells_column_labels(columns = everything())) %>%
  gt::tab_style(cell_text(style = "italic"), 
                locations = cells_body(columns = c(Genus, Species))) %>% 
  gt::tab_style(cell_borders(sides = "right", color = "grey50", weight = px(1.5)), 
                locations = cells_body(columns = Reference)) %>%
  gt::tab_style(cell_borders(sides = "right", color = "grey50", weight = px(1.5)), 
                locations = cells_column_labels(columns = Reference)) %>%
  gt::tab_footnote(footnote = "The species described in Kobayashi 2007.", 
                   locations = cells_column_labels(columns = Species)) %>%
  gt::tab_footnote(footnote = "The functional groups of the fungal species were inferred from previous studies.", 
                   locations = cells_column_labels(columns = Reference)) %>%
  gt::tab_footnote(footnote = "Species for which we have not been able to find any previous studies on their functional groups are classified as \"unknown\" functional groups.", 
                   locations = cells_column_labels(columns = `Functional group`)) %>%
  gt::tab_footnote(footnote = "The result of the FUNGuild analysis is shown.", 
                   locations = cells_column_labels(columns = c(Trait, 
                                                               Guild, 
                                                               `Trophic mode`, 
                                                               Confidence))) %>%
  opt_footnote_marks(marks = "standard") %>%
  tab_options(table.font.names = "Times New Roman")

show(path_gs_tbl)

gtsave(path_gs_tbl, file = paste0(path_out, "/231202_Sugi_phyllosphere_June_ITS/export/Pathogen_putative.png"), 
       zoom = 2, vwidth = 2835, vheight = 1417)
gtsave(path_gs_tbl, file = paste0(path_out, "/231202_Sugi_phyllosphere_June_ITS/export/Pathogen_putative.pdf"), 
       vwidth = 2835, vheight = 1417)
```

#envfit (not included in the paper published on Scientific Reports, 2025)

```{r}
#path_gs <- read.csv(paste0(path_out, "/231202_Sugi_phyllosphere_June_ITS/analysis/pathogen_Gs.csv"))
#path_gs <- column_to_rownames(path_gs, var = "X")


set.seed(56)
fit_path <- envfit(nmds_ITS$nmds, ps_ITS_path@otu_table)
fit_path$tax_table <- tax_table(ps_ITS_path)
show(fit_path)
df_fit_path <-data.frame(tax_table(ps_ITS_path)[rownames(fit_path$vectors$arrows), "Genus"], 
                         FuncType = path_gs$FuncType[sapply(rownames(fit_path$vectors$arrows), grep, path_gs$ASV)], 
                         fit_path$vectors$arrows,
                         r = fit_path$vectors$r,
                         p = fit_path$vectors$pvals)

save(fit_path, file = paste0(path_out, "/231202_Sugi_phyllosphere_June_ITS/analysis/envfit_bray_pathogen.Rdata"))
write.csv(df_fit_path, file = paste0(path_out, "/231202_Sugi_phyllosphere_June_ITS/analysis/envfit_bray_pathogen.csv"))
```

```{r}
load(paste0(path_out, "/231202_Sugi_phyllosphere_June_ITS/analysis/envfit_bray_pathogen.Rdata"))

path_score <- scores(fit_path, display = "vectors")
path_score <- rownames_to_column(data.frame(path_score))
path_score <- data.frame(path_score, 
                        Genus = fit_path$tax_table[path_score$rowname, "Genus"], 
                        Functype = path_gs$FuncType[sapply(path_score$rowname, grep, path_gs$ASV)], 
                        Functype2 = path_gs$FuncType2[sapply(path_score$rowname, grep, path_gs$ASV)], 
                        p = fit_path$vectors$pvals, 
                        arr_col = rep("grey", nrow(path_score)))

path_signif <- path_score[path_score$p < 0.1, ]
```

##plot envfit

```{r}
#filter population data only
popmeta <- metadata %>% 
  dplyr::select(population, fullname, latitude, longitude, altitude) %>% 
  group_by(population) %>% 
  summarize(population = first(population), 
            fullname = first(fullname), 
            latitude = first(latitude), 
            longitude = first(longitude), 
            altitude = first(altitude))

#assign numbers to populations in ascending order with respect to latitude
replace_rule <- 1:nrow(popmeta)
names(replace_rule) <- as.character(popmeta$population[sapply(sort(popmeta$latitude), 
                                                              grep, popmeta$latitude)])
```

```{r}
pop_num <- as.character(metadata$population)
pop_num <- unlist(lapply(pop_num, function(x) replace_rule[x]))
#set (site * population) levels
fac_sipop <- as.factor(paste0(metadata$site, LETTERS[pop_num]))
#assign colors to each element
#decrease transparency with increasing latitude
col_sipop <- c(alpha("blue", seq(0.2, 1, length.out = length(replace_rule))), 
               alpha("orange3", seq(0.2, 1, length.out = length(replace_rule)))
               )
```

```{r}
df_nmds_ITS <- data.frame(metadata, nmds_ITS$nmds$points)

p_ITS <- ggplot(data = df_nmds_ITS, 
                 mapping = aes(x = MDS1, y = MDS2, color = fac_sipop))
p_ITS <- p_ITS + theme_classic()
p_ITS <- p_ITS + coord_fixed(1)
p_ITS <- p_ITS + xlim(-2.5, 2) + ylim(-2, 2.5)
p_ITS <- p_ITS + scale_color_manual(values = col_sipop)
p_ITS <- p_ITS + geom_point()
p_ITS <- p_ITS + theme(legend.position = "none") +

#arrows to show the direction of Class abundance vectors
  geom_segment(data = path_signif,
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(2.5, "mm")), 
               color = "brown") +
  geom_text(data = path_signif,  
            aes(x = NMDS1, y = NMDS2, label = Genus), 
            color = "black") +
  #annotation to show the significance
  geom_segment(data = data.frame(x = 1.2, y = -1.8),
               aes(x = x, xend = (x + 0.2), y = y, yend = y),
               arrow = arrow(length = unit(2.5, "mm")), 
               color = "brown") +
  annotate("text", 
           x = 1.8, y = -1.8, 
           label = "P < 0.1")

#stress
p_ITS <- p_ITS + annotate("text", 
                          x = 1.5, y = -2,
                          label = paste0("Stress = ", 
                                         round(nmds_ITS$nmds$stress, digits = 4)))

show(p_ITS)
ggsave(filename = paste0(path_out, "/231202_Sugi_phyllosphere_June_ITS/figures/NMDS_bray_pathogen.png"), 
       plot = p_ITS, 
       dpi = 400, width = 150, height = 150, units = "mm")
```

#GLM
##each guild

```{r}
path_gs <- read.csv(paste0(path_out, "/231202_Sugi_phyllosphere_June_ITS/analysis/pathogen_Gs.csv"))
path_gs <- column_to_rownames(path_gs, var = "X")
FUNGuild_path <- read_tsv(paste0(path_out, "/231202_Sugi_phyllosphere_June_ITS/processed_data/FUNGuild/otu_table_Kobayashi.guilds.txt"))
path_gs_out <- read_tsv(file = paste0(path_out, "/231202_Sugi_phyllosphere_June_ITS/export/Pathogen_putative.tsv"))

load(paste0(path_out, "/231202_Sugi_phyllosphere_June_ITS/processed_data/phyloseq_pathogen.Rdata"))


abund_guild <- data.frame(ps_ITS_guild1@otu_table)
abund_guild2 <- data.frame(ps_ITS_guild2@otu_table)

abund_ga <- cbind(abund_guild, metadata) %>% 
  dplyr::select(c(ASV5543:ASV7510, site, latitude, PC1_clim, PC2_clim)) %>% 
  pivot_longer(!site:PC2_clim, names_to = "ASV", values_to = "abundance")

#fix the name of the guilds
guild <- rownames_to_column(data.frame(ps_ITS_guild1@tax_table[, 1]), var = "ASV")
guild$col <- c("brown", "blue3", "grey40", "tomato", "violetred", "pink")
guild$order <- c(6, 2, 5, 3, 4, 1)
#remove double space
guild$FuncType <- sub("  ", " ", guild$FuncType)
#capitalyze the first letter
guild$FuncType <- gsub("(^)([[:alpha:]])", "\\1\\U\\2", guild$FuncType, perl=TRUE)


abund_ga <- left_join(abund_ga, guild, by = "ASV")
#show "Unknown" at the last
abund_ga$FuncType <- factor(abund_ga$FuncType, 
                            levels = c(unique(abund_ga$FuncType)[-grep("Unknown", 
                                                                       unique(abund_ga$FuncType))], 
                                       "Unknown"))


ps_ITS_guild1@tax_table[, "FuncType"]
ps_ITS_guild2@tax_table[, "FuncType2"]
```

##mytheme

```{r}
#theme settings
mytheme <-theme(text = element_text(color = "black"), 
                axis.text = element_text(color = "black"),
                axis.title.y = element_text(size = 8), 
                
                legend.key.size = unit(9, "pt"), 
                legend.spacing.x = unit(0, "mm"))
```

##plot

```{r}
#show prediction and CI for the model with the total abundance tested on PC1_clim
path_total <- apply(ps_ITS_guild@otu_table, 1, sum)
model_p <- glm.nb(path_total~PC1_clim, data = metadata, link = log)
new_data <- data.frame(PC1_clim = seq(min(abund_ga$PC1_clim), 
                                      max(abund_ga$PC1_clim), 
                                      length.out = 100))
pred <- predict(model_p, newdata = new_data, type = "link", se.fit = TRUE)
crit <- qnorm(0.975)  # For 95% confidence intervals

new_data$fit <- exp(pred$fit)
new_data$lower <- exp(pred$fit - crit * pred$se.fit)
new_data$upper <- exp(pred$fit + crit * pred$se.fit)

#change the order of samples to show leaf rot fungi at the top
abund_plot <- abund_ga %>% arrange(order)

p_path <- ggplot(data = abund_plot, 
                 mapping = aes(x = PC1_clim, y = abundance, color = FuncType)) + 
  scale_color_manual(values = guild$col[sapply(levels(abund_ga$FuncType), grep, guild$FuncType)]) + 
  geom_line(data = new_data, aes(y = fit), linewidth = 0.7, color = "black") +
  geom_ribbon(data = new_data, 
              aes(ymin = lower, ymax = upper, y = fit),
              alpha = 0.05, linewidth = 0, color = NA, fill = "black") +
  geom_point() + 
  theme_classic() + mytheme +
  xlab("Climatic PC1") + ylab("Abundance of the putative pathogenic fungi after rarefaction") + labs(color = "Putative functional group") + NULL
  
show(p_path)

ggsave(filename = paste0(path_out, "/231202_Sugi_phyllosphere_June_ITS/figures/pathogen_climate1.png"), 
       plot = p_path, 
       dpi = 300, width = 160, height = 90, units = "mm")
#white, fruit rot fungi seemed to be abundant on saplings from warm region
```

```{r}
p_path <- ggplot(data = abund_ga, 
                 mapping = aes(x = PC2_clim, y = abundance, color = FuncType)) + 
  geom_point() + 
  scale_color_manual(values = guild$col[sapply(levels(abund_ga$FuncType), grep, guild$FuncType)]) + 
  theme_classic() + mytheme +
  xlab("Climatic PC2") + ylab("Abundance after rarefaction") + labs(color = "Guild", shape = "Site")
  

show(p_path)

ggsave(filename = paste0(path_out, "/231202_Sugi_phyllosphere_June_ITS/figures/pathogen_climate2.png"), 
       plot = p_path, 
       dpi = 300, width = 160, height = 90, units = "mm")
```

###GLM for each functional type

```{r}
#test relationships between pathogen abundance and climate or site
#exclude leaf rot pathogens (since they were absent on all saplings but one)
model_brown <- glm.nb(abund_guild[, "ASV5543"] ~ site + PC1_clim, 
                      data = metadata, method = "glm.fit", link = log)
##treat Xeromphalina as brown rot fungi
model_brown2 <- glm.nb(abund_guild2[, "ASV5543"] ~ site + PC1_clim, 
                      data = metadata, method = "glm.fit", link = log)

model_white <- glm.nb(abund_guild[, "ASV5249"] ~ site + PC1_clim, 
                      data = metadata, method = "glm.fit", link = log)

#pathogens with unknown guild were absent on all saplings at Tsukuba
model_unknown <- glm.nb(abund_guild[, "ASV2829"] ~ PC1_clim, 
                      data = metadata, method = "glm.fit", link = log)

#trunk rot pathogens were ansent on all saplings at Kawatabi
model_trunk <- glm.nb(abund_guild[, "ASV7543"] ~ PC1_clim, 
                      data = metadata, method = "glm.fit", link = log)

model_root <- glm.nb(abund_guild[, "ASV2762"] ~ site + PC1_clim, 
                      data = metadata, method = "glm.fit", link = log)

model_fruit <- glm.nb(abund_guild[, "ASV7510"] ~ site + PC1_clim, 
                      data = metadata, method = "glm.fit", link = log)
```

```{r}
summary <- list(brown = summary(model_brown), 
                brown2 = summary(model_brown2), 
                white = summary(model_white), 
                unknown = summary(model_unknown), 
                trunk = summary(model_trunk), 
                root = summary(model_root), 
                fruit = summary(model_fruit))
```

#total pathogen abundance

```{r}
path_total <- apply(ps_ITS_guild@otu_table, 1, sum)

plot(metadata$PC1_clim, path_total, col = metadata$site)

model_total <- glm.nb(path_total~site + PC1_clim + PC2_clim, 
                      data = metadata, link = log)
summ_tot <- summary(model_total)

#abundance of pathogens were associated with climate(negatively with warmth) at distributions of host cedar populations
```

#adjusted p value

```{r}
#bind up all test results
df_total <- cbind(pathogen = "all", 
                  call = paste(summ_tot[["call"]], collapse = ", "), 
                  summ_tot[["coefficients"]], 
                  aic = summ_tot[["aic"]])

df_list <- list()
for (i in 1:length(summary)) {
  df_list[[i]] <- cbind(pathogen = names(summary[i]), 
                      call = paste(summary[[i]][["call"]], collapse = ", "), 
                      summary[[i]][["coefficients"]], 
                      aic = summary[[i]][["aic"]])
  rownames(df_list[[i]]) <- paste0(rownames(df_list[[i]]), "_", 
                                   names(summary[i]))
}

df_res <- rbind(df_total, do.call("rbind", df_list))
```

```{r}
#calculate adjusted p value
#remove values for intercepts (which are not tested here)
df_res_adjust <- df_res[-grep("Intercept", rownames(df_res)), ]
df_res_adjust <- df_res_adjust[-grep("brown2", rownames(df_res_adjust)), ]

df_res_adjust <- cbind(df_res_adjust, 
                       p_adj = p.adjust(df_res_adjust[, "Pr(>|z|)"], method = "BH"))

#brown rot fungi were abundant at Kawatabi
#pathogens with unknown guild were absent on all saplings at Tsukuba
#trunk rot pathogens were absent on all saplings at Kawatabi

#save results
write.csv(df_res, file = paste0(path_out, "/231202_Sugi_phyllosphere_June_ITS/analysis/glm_pathogen.csv"))
write.csv(df_res_adjust, file = paste0(path_out, "/231202_Sugi_phyllosphere_June_ITS/analysis/glm_pathogen_adj.csv"))
```

